Best Practices for Alchemical Free Energy Calculations: v\versionnumber

### 2020-01-20 paper/manuscript.tex
Updates from Overleaf

Alchemical calculations can be used in a variety of scenarios, such as computing the free energy of a conformational change for a molecule with a high barrier to interconversion (Fig.~\ref{fig:fig_what_is_alchemy} (A)); 

@@

computing partition ($\log P$) or distribution ($\log D$) coefficients between environments (Fig.~\ref{fig:fig_what_is_alchemy} (B))~\cite{rustenburg2016measuring, bosisio2016blinded} or compartments or membranes (Fig.~\ref{fig:fig_what_is_alchemy} (C))~\cite{corey2019insights}. 
Furthermore, alchemical calculations are frequently used to estimate changes in free energies upon modifying a ligand or protein: a protein residue can be alchemically mutated to probe the impact on binding affinity (Fig.~\ref{fig:fig_what_is_alchemy} (D))\cite{hauser2018predicting,aldeghi2018accurate} or changes in protein thermostability~\cite{seeliger2010protein,gapsys2016insights,gapsys2016accurate,aldeghi2019accurate}; the entire ligand can be alchemical transferred from protein to solvent in an absolute binding free energies (Fig.~\ref{fig:fig_what_is_alchemy} (E))~\cite{mobley2007predicting,aldeghi2015accurate,aldeghi2017predictions}; and small alchemical modifications can be made between chemically related ligands to estimate differences in binding free energies (Fig.~\ref{fig:fig_what_is_alchemy} (F))~\cite{wang2015accurate,mey2016blinded,gapsys2020large}.

@@

    \caption{\textbf{Illustration of common types of free energies differences that can be calculated using alchemical free energy methods.} \textbf{A}: Change in free energy due to a conformational change of the molecule. \textbf{B}: Partition coefficient such as $\log P$ or $\log D$ depend on a change in free energy between different phases; here, as an example the partition coefficient between methanol and water is shown. \textbf{C}: Free energy difference associated with the insertion of a molecule into a membrane. \textbf{D]: Effect of mutations of protein or host residues on free energies of binding. \textbf{E}: Absolute free energies of binding of a small molecule to a host (e.g. protein), \textbf{F}: Relative free energy of binding of one molecule with respect to another, here toluene and benzyl alcohol.

@@

\todo[inline, color=green!20]{ASJSM: This section needs to be expanded and give a good introduction to the theoretical background without going into too much detail.}

@@

The equilibrium rate of binding is given by the ratio of concentrations of product $[PL]$ and reactants $[P]$, $[L]$:

@@

In order to define a reference state, the standard state is usually defined as $1 \mathrm{mol}/\mathrm{L}$ and $c^\circ$ is the standard state concentration. Thus, the binding free energy is given by:
    \Delta G_{\mathrm{bind}} = -RT\ln K_b^{\circ}.
One way to estimate the binding rate is by computing the probability of finding the molecular system in the bound state or the unbound state, making use of partition function ratios, i.e.

\begin{equation}
    \Delta G_{\mathrm{bind}} = -RT\ln\frac{\int_\mathrm{bound} \exp(-\beta U(\mathbf{x}))d\mathbf{x}}{\int_\mathrm{unbound} \exp(-\beta U(\mathbf{x}))d\mathbf{x}}
\end{equation}
Or more generally in a simplified form the free energy difference between the bound and unbound states is:
\begin{equation}
\Delta f \equiv f(\mathrm{bound}) - f(\mathrm{unbound}) = -\ln\frac{Z(\mathrm{unbound})}{Z({\mathrm{bound}})},
\end{equation}
where $Z = \int_\mathrm{bound} \exp(-\beta U(\mathbf{x}))d\mathbf{x}$ represents the configurational integral. 
How can we compute $\Delta G_{\mathrm{bind}}$ from a molecular dynamics simulation? Running a direct simulation of the binding events is too computationally costly, meaning we cannot easily estimate the configurational integrals of the bound and unbound state from the simulation. This is where alchemical free energy simulations come in. We can write down a thermodynamic cycle as seen in ~\ref{fig:fig_binding_thermodynamic_cycle}, for a relative free energy simulation. Based on this cycle we can compute the relative free energy of binding between two compounds $A$ and $B$, by going along the arrow of $\Delta G_{\mathrm{bound}}$, where we introduce a second compound that is similar to the first one. Now $\Delta G_{\mathrm{bound}}$ can be expressed as a ratio of partition functions of toluene (A) and benzyl alcohol (B). 
The defining characteristic of alchemical free energy calculations is the use of a series of alchemically-modified potential functions $U(x; \lambda)$ in which an alchemical parameter $\lambda$ modulates interactions in a manner that cannot occur in real chemical systems.
One or more simulations are used to collect data from a multitude of alchemical states to compute a free energy difference between a chemical state ($\lambda_0$) and another chemical or alchemical reference state ($\lambda_1$),
\begin{eqnarray}
\Delta f &\equiv& f(\lambda_1) - f(\lambda_0) = - \ln \frac{Z(\lambda_1)}{Z(\lambda_0)} , \label{equation:dimensionless-free-energy-difference}
\end{eqnarray}
where the dimensionless free energy $f(\lambda) \equiv \beta F(\lambda)$ is given in terms of partition functions $Z(\lambda)$,
\begin{eqnarray}
Z(\lambda) &=& \int dx \, e^{-u(x; \lambda)} .
\label{equation:partition-function-definition}
\end{eqnarray}
Here, the inverse thermal energy $\beta \equiv (k_B T)^{-1}$ where $k_B$ is the Boltzmann constant and $T$ is the absolute temperature, and the \emph{reduced potential} $u(x; \lambda)$~\cite{shirts2008statistically} is generally given by a trace over thermodynamic parameters with their conjugate dynamical variables,
\begin{eqnarray}
u(x;\lambda) &\equiv& \beta \left[ U(x;\lambda) + p \, V(x) + \sum_{i=1}^N \mu_i \, N_i(x) + \cdots \right] . \label{equation:reduced-potential}
\end{eqnarray}
Here, the collection of thermodynamic and alchemical parameters $\theta \equiv \{\beta, \lambda, p, \mu, \ldots\}$ defines a \emph{thermodynamic state}.
The physical transformation of interest---such as a protein-ligand association process, or transfer free energy among phases---may involve a thermodynamic cycle that requires several free energy differences $\Delta f$ to be computed in order to produce an estimate of the overall desired quantity.

\begin{figure}
    \includegraphics[width=0.95\linewidth]{figures/fig2_therm_cyc/Figure.pdf}
    \caption{{\bf Thermodynamic cycle for computing the relative free energy of binding ($\Delta \Delta G$) between two related small molecules to a supramolecular host.}
    The relative binding free energy difference between two small molecules, $\Delta \Delta G_{\mathrm{bind}, A \rightarrow B} \equiv \Delta G_{\mathrm{bind}, B} - \Delta G_{\mathrm{bind}, A}$---here benzyl alcohol (top) to toluene (bottom)---can be computed as a difference between two alchemical transformations, $\Delta G_\mathrm{bound} - \Delta G_\mathrm{solvated}$, where $\Delta G_\mathrm{bound}$ represents the free energy change of transforming $A \rightarrow B$ in complex and $\Delta G_\mathrm{solvated}$ the free energy change of transforming $A \rightarrow B$ in solvent.}
    \label{fig:fig_binding_thermodynamic_cycle}
\end{figure}

@@

    \caption{\textbf{Examples of perturbation networks} \textbf{A} Star shaped network with the crystal structure in the center. \textbf{B} Network with cycle closures (see more on this in Sec.~\ref{sec:are-they-good}). Arrows indicate the direction of the perturbation. Green cycles indicate good cycle closure, red poor cycle closure. The red arrow indicates a poorly converged simulation that would give rise to bad cycle closures. The diamond indicates the use of a crystal structure binding mode.}

@@

    \includegraphics[width=0.95\linewidth]{figures/fig5_thermo_cyc_abs/Figure.pdf}
    \caption{\textbf{Thermodynamic cycle required for an aboslute free energy calculation -- absolute free energy of binding example} The fully interacting ligand in water (\textbf{A}), has its charges turned off to pass to \textbf{B} followed by turning of van der Waals terms, resulting in a non-interacting ligand in water in \textbf{C}. Restraints are used on the fully interacting ligand in the binding site of a protein or host molecule \textbf{D}. The next step is to turn off the charges again (\textbf{E}) followed by the van der Waals interactions resulting in a non-interacting complex state (\textbf{F}). Free energyes can be computed as $\Delta G_{bind} = (\Delta G^{\mathrm{elec}}_{\mathrm{solv}}+ \Delta G^{\mathrm{VdW}}_{\mathrm{solv}})-(\Delta G^{\mathrm{elec}}_{\mathrm{bound}}+ \Delta G^{\mathrm{VdW}}_{\mathrm{bound}}$).

@@

    \caption{Alchemical intermediates are createtd by making the potential energy depend on a an additional variable $\lambda$ that interpolates between the chemical endpoints. In \textbf{A}, at $\lambda=0$ the molecule is a fully interacting phenol and at $\lambda=1$,  a fully interacting benzene.  \textbf{B} shows an illustration of the probability distribution of the potential energies as the switching function takes values of $\lambda=0$ to $\lambda=1$, intermediates are required for a sufficient overlap in potential energies to estimate a free energy difference between $\lambda=0$ and $\lambda=1$.
    One of the most efficient families of intermediate pathways  are softcore potentials, whose shape is illustrated in \textbf{C} with dependence on $\lambda$ and \textbf{D} with dependence on the potential energy. Note how when $\lambda$ approaches 0, the energy smoothly approaches zero at all $r$, a necessary requirement for efficient calculations.  }

@@

Another way of stating this is that intermediate states should sample molecular configurations that are as similar as possible to their neighboring states. The more similar the configurations are between intermediate states, the lower the statistical uncertainty is in the estimate of free energy between intervals. This can be proven directly from the BAR and MBAR formulas~\cite{bennett1976efficient,klimovich2015guidelines}, though the exact same principles apply for TI. For a 'good' path to work sufficient overlap between pontential energies is required. Fig.~\ref{fig:fig_what_is_lambda} \textbf{A} and \textb{B} try to illustrate this. Fig.~\ref{fig:fig_what_is_lambda} \textbf{A} shows in a pictorial way a soft-core potential can be applied across different $\lamda$s. Fig.~\ref{fig:fig_what_is_lambda} \textbf{B} illustrates the potential energy distributions at the different $\lambda$ intermediates, with sufficient overlap between neighboring $\lambda$ states to ensure that reweighting estimators such as MBAR can be used for analysis (see Sec.\ref{subsec:estimators}). The actual transformation is best handled with soft-core potentials of the form shown in Fig.\ref{fig:fig_what_is_lambda} \textbf{C} and \textbf{B}, with more details given below. 

@@

    U(\vec{r_{ij}},\lambda) = 4\epsilon_{ij} \lambda \left(\frac{1}{(\alpha(1-\lambda) + (r_{ij}/\sigma_{ij})^6)^2} +  \frac{1}{\alpha(1-\lambda) + (r_{ij}/\sigma_{ij})^6}\right)\label{eq:softcore}
Where $r_{ij}$ is the distance between two particles $i$ and $j$, $\epsilon_{ij}$ and $\sigma_{ij}$ are the Lennard-Jones parameters corresponding to the interaction between particles $i$ and $j$, and $\alpha$ is a constant (0.5 is optimal for the functional form shown above).  This functional form has exactly the property we are looking for: it recovers the Lennard-Jones potential when $\lambda=1$, and the other endpoint ($\lambda=0$), it is exactly zero for all $r_{ij}$ everywhere, and as $\lambda$ goes to zero, the  $\alpha(1-\lambda)$ term lowers the infinite energy in the core.  There are several different variants of the same functional form~\cite{zacharias1994separationshifted, beutler1994avoiding,pham2011identifying}, but the one in given in eq.~\ref{eq:softcore} is easy to understand and implement and fairly numerically stable. This functional form is shown in parts \textbf{C} and \textbf{D} of Fig.~\ref{fig:fig_what_is_lambda}.
It has been shown that more complicated forms are not significantly more efficient than eq.~\cite{pham2012optimal}.  We recommend using the softcore potential given in eq.~\cite{eq:softcore}, unless there is a compelling reason otherwise. Using a similar equation to eq.~\cite{eq:softcore} may be acceptable in most circumstances if that is what is supported in your chosen software. However, if you are inserting or removing entire atomic sites, we heavily recommend against using the linear approach; it will be very difficult to get correct results. 

>>>>

Alchemical calculations can be used in a variety of scenarios, such as computing the free energy of a conformational change for a molecule with a high barrier to interconversion (Fig.~\ref{fig:fig_what_is_alchemy} (\textbf{A})); 

@@

computing partition ($\log P$) or distribution ($\log D$) coefficients between environments (Fig.~\ref{fig:fig_what_is_alchemy} (\texbf{B}))~\cite{rustenburg2016measuring, bosisio2016blinded} or compartments or membranes (Fig.~\ref{fig:fig_what_is_alchemy} (\textbf{C}))~\cite{corey2019insights}. 
Furthermore, alchemical calculations are frequently used to estimate changes in free energies upon modifying a ligand or protein: a protein residue can be alchemically mutated to probe the impact on binding affinity (Fig.~\ref{fig:fig_what_is_alchemy} (\textbf{D}))\cite{hauser2018predicting,aldeghi2018accurate} or changes in protein thermostability~\cite{seeliger2010protein,gapsys2016insights,gapsys2016accurate,aldeghi2019accurate}; the entire ligand can be alchemical transferred from protein to solvent in an absolute binding free energies (Fig.~\ref{fig:fig_what_is_alchemy} (\textbf{E}))~\cite{mobley2007predicting,aldeghi2015accurate,aldeghi2017predictions}; and small alchemical modifications can be made between chemically related ligands to estimate differences in binding free energies (Fig.~\ref{fig:fig_what_is_alchemy} (F))~\cite{wang2015accurate,mey2016blinded,gapsys2020large}.

@@

    \caption{\textbf{Illustration of common types of free energies differences that can be calculated using alchemical free energy methods.} \textbf{A}: Change in free energy due to a conformational change of the molecule. \textbf{B}: Partition coefficient such as $\log P$ or $\log D$ depend on a change in free energy between different phases; here, as an example the partition coefficient between methanol and water is shown. \textbf{C}: Free energy difference associated with the insertion of a molecule into a membrane. \textbf{D}: Effect of mutations of protein or host residues on free energies of binding. \textbf{E}: Absolute free energies of binding of a small molecule to a host (e.g. protein), \textbf{F}: Relative free energy of binding of one molecule with respect to another, here toluene and benzyl alcohol.

@@

@@

The equilibrium constant  ($K_b^{\circ}$) is given by the ratio of concentrations of product $[PL]$ and reactants $[P]$, $[L]$:

@@

In order to define a reference state, the standard state is usually defined as $1 \mathrm{mol}/\mathrm{L}$ and $c^\circ$ is the standard state concentration. Thus, the Gibbs free energy of binding $\Delta G_{\mathrm{bind}}$ is given by:
    \Delta G_{\mathrm{bind}} = -k_BT\ln K_b^{\circ},
where $k_B$ is the Boltzmann constant and $T$ the temperature of the system. 
One way to estimate the equilibrium constant is by computing the probability of finding the molecular system in the bound state ($PL$) or the unbound state ($P+L$), making use of partition function ratios, and assuming that the contribution of water compressibility at 1 atm has a negligible effect on the free energy of binding~\cite{shirts2012best, gilson1997statisticalthermodynamic} i.e.
    \Delta G_{\mathrm{bind}} \cong \Delta F= -k_BT\ln\frac{\int_\mathrm{bound} \exp(-\beta U(\mathbf{x}))d\mathbf{x}}{\int_\mathrm{unbound} \exp(-\beta U(\mathbf{x}))d\mathbf{x}},
where $\beta = \frac{1}{k_BT}$ and $F$ is the Helmholtz free energy. 
Or more generally in a simplified form the free energy difference ($\Delta f$) between the bound and unbound states is:
\Delta f \equiv f(\mathrm{bound}) - f(\mathrm{unbound}) = -\ln\frac{Z(\mathrm{unbound})}{Z({\mathrm{bound}})},
where $Z = \int_\mathrm{bound} \exp(-\beta U(\mathbf{x}))d\mathbf{x}$ represents the configurational integral. 
Unfortuantely it is not possible to evaluate the integral required for the configurational integral exactly, but can be approximated using molecular dynamics simulations. 
So, how can we compute $\Delta G_{\mathrm{bind}}$ from a molecular dynamics simulation? Running a direct simulation of the binding events is too computationally costly, meaning we cannot easily estimate the configurational integrals of the bound and unbound state from the simulation, as typically binding events happen at a $\mu$s scale, and simulations reach around 1 $\mu$s/day. For a reliable estimate of a single free energy of binding many days of simulations would be required, making these type of calculations uninteresting at a pharamaceutical timescale. Instead, we can run alchemical free energy simulations, by making use of the fact that free energy is a state function. We can write down a thermodynamic cycle as seen in ~\ref{fig:fig_binding_thermodynamic_cycle}, for a relative free energy simulation. Based on this cycle we can compute the relative free energy of binding between two compounds $A$ and $B$, by going along the arrow of $\Delta G_{\mathrm{bound}}$, where we introduce a second compound that is similar to the first one. Now $\Delta G_{\mathrm{bound}}$ can be expressed as a ratio of partition functions of toluene (A) and benzyl alcohol (B). 
The defining characteristic of alchemical free energy calculations is the use of a series of alchemically-modified potential functions $U(x; \lambda)$ in which an alchemical parameter $\lambda$ modulates interactions in a manner that cannot occur in real chemical systems.
One or more simulations are used to collect data from a multitude of alchemical states to compute a free energy difference between a chemical state ($\lambda_0$) and another chemical or alchemical reference state ($\lambda_1$),
\Delta f &\equiv& f(\lambda_1) - f(\lambda_0) = - \ln \frac{Z(\lambda_1)}{Z(\lambda_0)} , \label{equation:dimensionless-free-energy-difference}
where the dimensionless free energy $f(\lambda) \equiv \beta F(\lambda)$ is given in terms of partition functions $Z(\lambda)$,
Z(\lambda) &=& \int_{\Gamma} d\vec{q} \, e^{-u(\vec{q}; \lambda)} ,
\label{equation:partition-function-definition}
, where $\Gamma$ signifies the phase space volume over which to integrate. 
The \emph{reduced potential} $u(\vec{q}; \lambda)$~\cite{shirts2008statisticallya} is generally given by a trace over thermodynamic parameters with their conjugate dynamical variables,
u(\vec{q};\lambda) &\equiv& \beta \left[ U(\vec{q};\lambda) + p \, V(\vec{q}) + \sum_{i=1}^N \mu_i \, N_i(\vec{q}) + \cdots \right] . \label{equation:reduced-potential}
Here, the collection of thermodynamic and alchemical parameters $\theta \equiv \{\beta, \lambda, p, \mu, \ldots\}$ defines a \emph{thermodynamic state}.
The physical transformation of interest---such as a protein-ligand association process, or transfer free energy among phases---may involve a thermodynamic cycle that requires several free energy differences $\Delta f$ to be computed in order to produce an estimate of the overall desired quantity, see for example the thermodynamic cycle for an absolute free energy calculation as seen in~\ref{fig:fig_absolute_thermodynamic_cycle}.

\begin{figure}
    \includegraphics[width=0.95\linewidth]{figures/fig2_therm_cyc/Figure.pdf}
    \caption{{\bf Thermodynamic cycle for computing the relative free energy of binding ($\Delta \Delta G$) between two related small molecules to a supramolecular host.}
    The relative binding free energy difference between two small molecules, $\Delta \Delta G_{\mathrm{bind}, A \rightarrow B} \equiv \Delta G_{\mathrm{bind}, B} - \Delta G_{\mathrm{bind}, A}$---here benzyl alcohol (top) to toluene (bottom)---can be computed as a difference between two alchemical transformations, $\Delta G_\mathrm{bound} - \Delta G_\mathrm{solvated}$, where $\Delta G_\mathrm{bound}$ represents the free energy change of transforming $A \rightarrow B$ in complex, i.e. bound to a host molecule, and $\Delta G_\mathrm{solvated}$ the free energy change of transforming $A \rightarrow B$ in solvent, typically water.}
    \label{fig:fig_binding_thermodynamic_cycle}
\end{figure}

@@

    \caption{\textbf{Examples of perturbation networks} (\textbf{A}) Star shaped network with the crystal structure in the center. (\textbf{B}) Network with cycle closures (see more on this in Sec.~\ref{sec:are-they-good}). Arrows indicate the direction of the perturbation. Green cycles indicate good cycle closure, red poor cycle closure. The red arrow indicates a poorly converged simulation that would give rise to bad cycle closures. The diamond indicates the use of a crystal structure binding mode.}

@@

    \includegraphics[width=0.95\linewidth]{paper/figures/fig5_thermo_cyc_abs/Figure.pdf}
    \caption{\textbf{Thermodynamic cycle required for an aboslute free energy calculation -- absolute free energy of binding example} The fully interacting ligand in water (\textbf{A}), has its charges turned off to pass to (\textbf{B}) followed by turning of van der Waals terms, resulting in a non-interacting ligand in water in (\textbf{C}). Restraints are used on the fully interacting ligand in the binding site of a protein or host molecule (\textbf{D}). The next step is to turn off the charges again (\textbf{E}) followed by the van der Waals interactions resulting in a non-interacting complex state (\textbf{F}). Free energyes can be computed as $\Delta G_{bind} = (\Delta G^{\mathrm{elec}}_{\mathrm{solv}}+ \Delta G^{\mathrm{VdW}}_{\mathrm{solv}})-(\Delta G^{\mathrm{elec}}_{\mathrm{bound}}+ \Delta G^{\mathrm{VdW}}_{\mathrm{bound}}$).

@@

    \caption{Alchemical intermediates are createtd by making the potential energy depend on a an additional variable $\lambda$ that interpolates between the chemical endpoints. In (\textbf{A}), at $\lambda=0$ the molecule is a fully interacting phenol and at $\lambda=1$,  a fully interacting benzene. (\textbf{B}) shows an illustration of the probability distribution of the potential energies as the switching function takes values of $\lambda=0$ to $\lambda=1$, intermediates are required for a sufficient overlap in potential energies to estimate a free energy difference between $\lambda=0$ and $\lambda=1$.
    One of the most efficient families of intermediate pathways  are softcore potentials, whose shape is illustrated in (\textbf{C}) with dependence on $\lambda$ and (\textbf{D}) with dependence on the potential energy. Note how when $\lambda$ approaches 0, the energy smoothly approaches zero at all $r$, a necessary requirement for efficient calculations.  }

@@

Another way of stating this is that intermediate states should sample molecular configurations that are as similar as possible to their neighboring states. The more similar the configurations are between intermediate states, the lower the statistical uncertainty is in the estimate of free energy between intervals. This can be proven directly from the BAR and MBAR formulas~\cite{bennett1976efficient,klimovich2015guidelines}, though the exact same principles apply for TI. For a 'good' path to work sufficient overlap between pontential energies is required. Fig.~\ref{fig:fig_what_is_lambda} (\textbf{A}) and (\textb{B}) try to illustrate this. Fig.~\ref{fig:fig_what_is_lambda} (\textbf{A}) shows in a pictorial way a soft-core potential can be applied across different $\lamda$s. Fig.~\ref{fig:fig_what_is_lambda} (\textbf{B}) illustrates the potential energy distributions at the different $\lambda$ intermediates, with sufficient overlap between neighboring $\lambda$ states to ensure that reweighting estimators such as MBAR can be used for analysis (see Sec.\ref{subsec:estimators}). The actual transformation is best handled with soft-core potentials of the form shown in Fig.\ref{fig:fig_what_is_lambda} (\textbf{C}) and (\textbf{B}), with more details given below. 

@@

    U(\vec{r_{ij}},\lambda) = 4\epsilon_{ij} \lambda \left(\frac{1}{(\alpha(1-\lambda) + (r_{ij}/\sigma_{ij})^6)^2} +  \frac{1}{\alpha(1-\lambda) + (r_{ij}/\sigma_{ij})^6}\right)
    \label{eq:softcore}
Where $r_{ij}$ is the distance between two particles $i$ and $j$, $\epsilon_{ij}$ and $\sigma_{ij}$ are the Lennard-Jones parameters corresponding to the interaction between particles $i$ and $j$, and $\alpha$ is a constant (0.5 is optimal for the functional form shown above).  This functional form has exactly the property we are looking for: it recovers the Lennard-Jones potential when $\lambda=1$, and the other endpoint ($\lambda=0$), it is exactly zero for all $r_{ij}$ everywhere, and as $\lambda$ goes to zero, the  $\alpha(1-\lambda)$ term lowers the infinite energy in the core.  There are several different variants of the same functional form~\cite{zacharias1994separationshifted, beutler1994avoiding,pham2011identifying}, but the one in given in eq.~\ref{eq:softcore} is easy to understand and implement and fairly numerically stable. This functional form is shown in parts (\textbf{C}) and (\textbf{D}) of Fig.~\ref{fig:fig_what_is_lambda}.
It has been shown that more complicated forms are not significantly more efficient than eq.~\cite{pham2012optimal}.  We recommend using the softcore potential given in eq.~\ref{eq:softcore}, unless there is a compelling reason otherwise. Using a similar equation to eq.~\ref{eq:softcore} may be acceptable in most circumstances if that is what is supported in your chosen software. However, if you are inserting or removing entire atomic sites, we heavily recommend against using the linear approach; it will be very difficult to get correct results. 
